### 2012-02-04 contents.tex
Version 1.

However, mining bug repositories has its own risks. Previous research has identified problems of missing data (e.g., rationale, traceability links between reported bug fixes and source code changes) and inaccurate data (e.g., misclassification of bugs) \cite{Aranda2009}. %,Bird2009,Nguyen2010,Ayari2007}.

@@

	In order to perform the desired analyses, we needed access to the data recorded by Bugzilla for a specific project, including status changes and comments. We have found such data for two projects---Eclipse and NetBeans---from the domain of integrated development environments. The data were made available as part of the Mining Software Repositories 2011 Challenge\footnote{\url{http://www.msrconf.org/msr-challenge.html}} in the form of MySQL database dumps. The files contain all data from the respective databases, except for developer profiles, that were omitted for privacy reasons.

@@

	NetBeans\footnote{\url{http://netbeans.org/about/history.html}} started  as a student project in 1996. It was bought by Sun Microsystems in October, 1999, and it was open sourced in March, 2000. The data set contains 185,578 bug reports for its 39 subprojects, from June, 1998 to June, 2010.

@@

	Four subprojects were chosen for further analysis: Eclipse/Platform, Eclipse/EMF, Netbeans/Platform, and Netbeans/VersionControl. The Platform subprojects are the main subprojects for the respective IDEs, so they are both important and representative of each projects' philosophy.

@@

	% In order to obtain insight about the research questions, we read a sample of bug reports from both projects. For each project, two subprojects were chosen randomly among the projects with most verified bugs, resulting in the following subprojects: Eclipse/Platform, Eclipse/EMF, Netbeans/platform, Netbeans/versioncontrol. We have sampled 20 bug reports for each subproject, totalizing 80 bug reports.
	%

@@

	In order to discover the verification techniques used by the subprojects, we have looked at the comments written by developers when they mark a bug as \VERIFIED\ (meaning that the fix was accepted) or \REOPENED\ (meaning that the fix was rejected). First, an informal analysis was performed, by reading a sample of such comments, looking for textual patterns that indicate that a particular technique was used. Then, based on the informal analysis and on previous knowledge, a regular expression was built for each technique, that should match comments in which there is evidence that that particular technique was used. The regular expressions were refined by looking at the results matched by them and modifying them to avoid false positives.

	The regular expressions can be downloaded at \url{http://...}. For the sake of simplicity, the following list describes only some substrings matched by the regular expressions for each technique:

@@

	During the informal analysis, it was observed that ad hoc testing is reported in many distinct ways that are difficult to capture by means of regular expressions. Because of this, the regular expression for is expected to match only a few comments.

@@

  Most verification comments do not provide any clue about the verification technique used. Most often, developers just state that the bug fix was verified, sometimes informing the build used to verify the fix.

@@

	\begin{table}
		\begin{center}
		\caption{Verification techniques for all four subprojects.}
		\begin{tabular}{l|r|r|r}
			% \toprule
			\textbf{\centering Project} & 
			\textbf{\centering Testing} & 
			\textbf{\centering Inspection} & 
			\textbf{\centering Ad Hoc} & 
			\hline
			Eclipse/Platform        & 260 (1.1\%) & 524 (2.2\%) & 74 (0.3\%) \\
			Eclipse/EMF             & 27 (0.7\%) & 2 (0.1\%) & 0 (0.0\%) \\
			NetBeans/Platform       & 94 (0.6\%) & 5 (0.0\%) & 0 (0.0\%) \\
			NetBeans/VersionControl & 4 (0.1\%) & 1 (0.0\%) & 0 (0.0\%) 
			% \bottomrule
		\end{tabular}
		\end{center}

@@

	While sorting through comments, we have found interesting information regarding the software verification process. In Eclipse/Platform, the developer who fixes a bug often someone else to verify the fix, by asking ``please verify, [developer name]''. If the bug is reopened, then the fixer and the verifier exchange roles. This behavior illustrates a structured bug fixing/verification process.

@@

	Conclusion: it's hard to determine the technique used by just looking at the data. Eclipse/Platform documents it better, sometimes.

@@

	Figure \ref{fig:when} shows plots of total accumulated number verifications over time for Eclipse/Platform (left) and NetBeans/Platform (right). Releases are plotted as dashed vertical lines.

@@

	For Eclipse/Platform, the graph shows the period between releases 3.5 and 3.6, including milestone releases every 6 to 7 weeks. It is clear from the graph that the verification activity is intensified in a few days preceding a milestone, represented in the graph by steeper ascents before the vertical lines. This pattern is an indicator of a verification phase. No such pattern was found in the other subprojects by analyzing their graphs (not shown here for brevity).

@@

	% 
	% We found evidence of a verification phase for Eclipse/Platform. In this project, verifications appear to be performed in bursts, with periods of intense activity separated by about 6 to 7 weeks of low or no activity. By taking a closer look in the period between the releases 3.5 and 3.6 (Figure XXX), it is clear that the periods of intense verification activity occur just before the milestones leading to 3.6.
	% 
	% No such pattern was found for Netbeans' subprojects.

@@

	% The first thing that comes to attention are ``cliff walls'', points in time when a large amount of bugs are verified, represented by steep, almost vertical portions of the plotted line. They appear in Eclipse/EMF and Netbeans/Platform. In Netbeans/Platform, all verifications in the cliff wall preceding version 5.5 share the same message: ``Verification of old issues''. For Eclipse/EMF, the message is ``Move to verified as per bug 206558'', which refers to the task of changing bugs marked as \FIXED\ to \VERIFIED.
	% 
	% Cliff walls represent a change in the development process, and the mass verification is needed in order to leverage useful features from Bugzilla (e.g., looking for verified bugs in order to aid the writing of release notes).

@@

	In NetBeans (both Platform and VersionControl) it is possible to infer the existence of a QA team, composed by approximately 20\% of the developers, that performs about 90\% of all verifications. In the Eclipse subprojects, there is no evidence of a dedicated QA team. In Eclipse/EMF no developer focus specifically on verification. In Eclipse/Platform, 2.4\% of the developers focus on verification tasks, contributing only to 1.1\% of the verifications.

@@

\begin{section}{Conclusion} \label{sec:conclusion}
	By analyzing four subprojects (two from Eclipse, two from NetBeans), we have found, using only data from bug repositories, subprojects with and without QA teams, with and without a well-defined verification phase. We also have found evidence of the application of automated testing and source code inspection, although such evidence was less reliable. Also, we have found cases in which marking a bug as \VERIFIED\ does not imply that any kind of software verification was actually performed.

	With the knowledge obtained from this exploratory research, we aim to improve and extend our previous work on the impact of independent verification on software quality. We can investigate, for example, whether verification performed by QA team is better than verification performed by other developers.

>>>>

However, mining bug repositories has its own risks. Previous research has identified problems of missing data (e.g., rationale, traceability links between reported bug fixes and source code changes) and inaccurate data (e.g., misclassification of bugs) \cite{Aranda2009,Bird2009}. %,Bird2009,Nguyen2010,Ayari2007}.

@@

	In order to perform the desired analyses, we needed access to the data recorded by Bugzilla for a specific project, including status changes and comments. We have found such data for two projects---Eclipse and NetBeans---from the domain of integrated development environments. The data were made available as part of the Mining Software Repositories 2011 Challenge\footnote{\url{http://www.msrconf.org/msr-challenge.html}} in the form of MySQL database dumps. The files contain all data from the respective databases, except for developer profiles, omitted for privacy reasons.

@@

	NetBeans\footnote{\url{http://netbeans.org/about/history.html}} started  as a student project in 1996. It was then bought by Sun Microsystems in October, 1999, and open sourced in March, 2000. The data set contains 185,578 bug reports for its 39 subprojects, from June, 1998 to June, 2010.

@@

	Four subprojects were chosen for further analysis: Eclipse/Platform, Eclipse/EMF, NetBeans/Platform, and NetBeans/VersionControl. The Platform subprojects are the main subprojects for the respective IDEs, so they are both important and representative of each projects' philosophy.

@@

	% In order to obtain insight about the research questions, we read a sample of bug reports from both projects. For each project, two subprojects were chosen randomly among the projects with most verified bugs, resulting in the following subprojects: Eclipse/Platform, Eclipse/EMF, NetBeans/platform, NetBeans/versioncontrol. We have sampled 20 bug reports for each subproject, totalizing 80 bug reports.
	%

@@

	In order to discover the verification techniques used by the subprojects, we have looked at the comments written by developers when they mark a bug as \VERIFIED\ (meaning that the fix was accepted) or \REOPENED\ (meaning that the fix was rejected). First, an informal analysis was performed, by reading a sample of such comments, looking for textual patterns that indicate that a particular technique was used. Then, based on the informal analysis and on previous knowledge, a regular expression was built for each technique, to match comments in which there is evidence that that particular technique was used. The regular expressions were refined by looking at the results matched by them and modifying them to avoid false positives.

	The regular expressions are available at \url{http://...}. For the sake of simplicity, the following list describes only some substrings matched by the regular expressions for each technique:

@@

	During the informal analysis, it was observed that ad hoc testing is reported in many distinct ways that are difficult to capture by means of regular expressions. Hence, its regular expression is expected to match only a few comments.

@@

@@

	\begin{table}
		\begin{center}
		\caption{Verification techniques for all four subprojects.}
		\begin{tabular}{l|r|r|r}
			% \toprule
			\textbf{\centering Project} & 
			\textbf{\centering Testing} & 
			\textbf{\centering Inspection} & 
			\textbf{\centering Ad Hoc} \\
			\hline
			Eclipse/Platform        & 260 (1.1\%) & 524 (2.2\%) & 74 (0.3\%) \\
			Eclipse/EMF             & 27 (0.7\%) & 2 (0.1\%) & 0 (0.0\%) \\
			NetBeans/Platform       & 94 (0.6\%) & 5 (0.0\%) & 0 (0.0\%) \\
			NetBeans/VersionControl & 4 (0.1\%) & 1 (0.0\%) & 0 (0.0\%) 
			% \bottomrule
		\end{tabular}
		\end{center}

@@

	While sorting through comments, we have found interesting information regarding the software verification process. Most verification comments do not provide any clue about the verification technique used. Most often, developers just state that the bug fix was verified, sometimes informing the build used to verify the fix.

	In Eclipse/Platform, the developer who fixes a bug often someone else to verify the fix, by asking ``please verify, [developer name]''. If the bug is reopened, then the fixer and the verifier exchange roles. This behavior illustrates a structured bug fixing/verification process.

@@

	% Conclusion: it's hard to determine the technique used by just looking at the data. Eclipse/Platform documents it better, sometimes.

@@

	Figure \ref{fig:when} shows plots of total accumulated number of verifications over time for Eclipse/Platform (left) and NetBeans/Platform (right). Releases are plotted as dashed vertical lines.

@@

	For Eclipse/Platform, the graph shows the period between releases 3.5 and 3.6, including milestone releases every 6 or 7 weeks. It is clear from the graph that the verification activity is intensified in a few days preceding a milestone, represented in the graph by steeper ascents before the vertical lines. This pattern is an indicator of a verification phase. No such pattern was found in the other subprojects by analyzing their graphs (not shown here for brevity).

@@

	% 
	% We found evidence of a verification phase for Eclipse/Platform. In this project, verifications appear to be performed in bursts, with periods of intense activity separated by about 6 to 7 weeks of low or no activity. By taking a closer look in the period between the releases 3.5 and 3.6 (Figure XXX), it is clear that the periods of intense verification activity occur just before the milestones leading to 3.6.
	% 
	% No such pattern was found for NetBeans' subprojects.

@@

	% The first thing that comes to attention are ``cliff walls'', points in time when a large amount of bugs are verified, represented by steep, almost vertical portions of the plotted line. They appear in Eclipse/EMF and NetBeans/Platform. In NetBeans/Platform, all verifications in the cliff wall preceding version 5.5 share the same message: ``Verification of old issues''. For Eclipse/EMF, the message is ``Move to verified as per bug 206558'', which refers to the task of changing bugs marked as \FIXED\ to \VERIFIED.
	% 
	% Cliff walls represent a change in the development process, and the mass verification is needed in order to leverage useful features from Bugzilla (e.g., looking for verified bugs in order to aid the writing of release notes).

@@

	In both NetBeans subprojects it is possible to infer the existence of a QA team, composed by approximately 20\% of the developers, that performs about 90\% of all verifications. In the Eclipse subprojects, there is no evidence of a dedicated QA team. In Eclipse/EMF no developer focus specifically on verification. In Eclipse/Platform, 2.4\% of the developers focus on verification tasks, contributing only to 1.1\% of the verifications.

@@

\begin{section}{Conclusion} \label{sec:conclusion}
	By analyzing four subprojects (two from Eclipse, two from NetBeans), we have found, using only data from bug repositories, subprojects with and without QA teams, with and without a well-defined verification phase. We also have found weaker evidence of the application of automated testing and source code inspection. Also, we have found cases in which marking a bug as \VERIFIED\ does not imply that any kind of software verification was actually performed.

	With the knowledge obtained from this exploratory research, we aim to improve and extend our previous work on the impact of independent verification on software quality. We can investigate, for example, whether verification performed by QA team is more effective than verification performed by other developers.
